---
title: "model_selection"
author: "Eren Kafadar"
date: "2024-11-26"
output: html_document
editor_options: 
  chunk_output_type: console
---
#Main Setup
```{r setup, include=FALSE}
diagnostics <- TRUE
generate <- FALSE
save_fits <- FALSE
code_path <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/ABCD_premie_code/"
fits_path <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/"
data_path <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/CSV/process_tables_5.1_2.0/"
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(dplyr)
library(magrittr)
library(readr)
library(stringr)
library(ggplot2)
library(gamlss)
require(gamlss.ggplots)
library(tidyr)
library(gamlssTools)
library(gridExtra)
library(ggseg)
library(purrr)

source(paste0(code_path, "inspect_model_fits.R"))
source(paste0(code_path, "generate_gamlss_model_fct.R"))
source(paste0(code_path, "load_data_fct.R"))

source(paste0(code_path, "pred_og_centile_EK.R"))
source(paste0(code_path, "wp_taki_EK.R"))

phenotypes <- readRDS(paste0(code_path, "phenotype_set.rds"))

#df to map variable names onto names displayed on figures
name_mapping <- data.frame(
  abbreviation = c(
     "smri_vol_cdk_totallh", "smri_vol_cdk_totalrh", "smri_vol_cdk_total",   
     "totalWM_cb", "totalWM_crb", "totalGM_crb",
     
    "smri_vol_scs_cbwmatterlh", "smri_vol_scs_ltventriclelh", "smri_vol_scs_inflatventlh",
    "smri_vol_scs_crbwmatterlh", "smri_vol_scs_crbcortexlh", "smri_vol_scs_tplh",
    "smri_vol_scs_caudatelh", "smri_vol_scs_putamenlh", "smri_vol_scs_pallidumlh",
    "smri_vol_scs_3rdventricle", "smri_vol_scs_4thventricle", "smri_vol_scs_bstem",
    "smri_vol_scs_hpuslh", "smri_vol_scs_amygdalalh", "smri_vol_scs_csf",
    "smri_vol_scs_lesionlh", "smri_vol_scs_aal", "smri_vol_scs_vedclh",
    
    "smri_vol_scs_cbwmatterrh", "smri_vol_scs_ltventriclerh", "smri_vol_scs_inflatventrh",
    "smri_vol_scs_crbwmatterrh", "smri_vol_scs_crbcortexrh", "smri_vol_scs_tprh",
    "smri_vol_scs_caudaterh", "smri_vol_scs_putamenrh", "smri_vol_scs_pallidumrh",
    "smri_vol_scs_hpusrh", "smri_vol_scs_amygdalarh", "smri_vol_scs_lesionrh",
    "smri_vol_scs_aar", "smri_vol_scs_vedcrh", 
    
    "smri_vol_scs_wmhint",
    "smri_vol_scs_wmhintlh", "smri_vol_scs_wmhintrh", 
    
    "smri_vol_scs_ccps",
    "smri_vol_scs_ccmidps", "smri_vol_scs_ccct", "smri_vol_scs_ccmidat",
    "smri_vol_scs_ccat", 
    
    "smri_vol_scs_wholeb", "smri_vol_scs_latventricles",
    "smri_vol_scs_allventricles", "smri_vol_scs_intracranialv", "smri_vol_scs_suprateialv",
    "smri_vol_scs_subcorticalgv",
    
    "smri_vol_cdk_banksstslh", "smri_vol_cdk_cdacatelh", "smri_vol_cdk_cdmdfrlh",
    "smri_vol_cdk_cuneuslh", "smri_vol_cdk_ehinallh", "smri_vol_cdk_fusiformlh",
    "smri_vol_cdk_ifpllh", "smri_vol_cdk_iftmlh", "smri_vol_cdk_ihcatelh",
    "smri_vol_cdk_locclh", "smri_vol_cdk_lobfrlh", "smri_vol_cdk_linguallh",
    "smri_vol_cdk_mobfrlh", "smri_vol_cdk_mdtmlh", "smri_vol_cdk_parahpallh",
    "smri_vol_cdk_paracnlh", "smri_vol_cdk_parsopclh", "smri_vol_cdk_parsobislh",
    "smri_vol_cdk_parstgrislh", "smri_vol_cdk_pericclh", "smri_vol_cdk_postcnlh",
    "smri_vol_cdk_ptcatelh", "smri_vol_cdk_precnlh", "smri_vol_cdk_pclh",
    "smri_vol_cdk_rracatelh", "smri_vol_cdk_rrmdfrlh", "smri_vol_cdk_sufrlh",
    "smri_vol_cdk_supllh", "smri_vol_cdk_sutmlh", "smri_vol_cdk_smlh",
    "smri_vol_cdk_frpolelh", "smri_vol_cdk_tmpolelh", "smri_vol_cdk_trvtmlh",
    "smri_vol_cdk_insulalh",
    
     "smri_vol_cdk_banksstsrh", "smri_vol_cdk_cdacaterh", "smri_vol_cdk_cdmdfrrh",
    "smri_vol_cdk_cuneusrh", "smri_vol_cdk_ehinalrh", "smri_vol_cdk_fusiformrh",
    "smri_vol_cdk_ifplrh", "smri_vol_cdk_iftmrh", "smri_vol_cdk_ihcaterh",
    "smri_vol_cdk_loccrh", "smri_vol_cdk_lobfrrh", "smri_vol_cdk_lingualrh",
    "smri_vol_cdk_mobfrrh", "smri_vol_cdk_mdtmrh", "smri_vol_cdk_parahpalrh",
    "smri_vol_cdk_paracnrh", "smri_vol_cdk_parsopcrh", "smri_vol_cdk_parsobisrh",
    "smri_vol_cdk_parstgrisrh", "smri_vol_cdk_periccrh", "smri_vol_cdk_postcnrh",
    "smri_vol_cdk_ptcaterh", "smri_vol_cdk_precnrh", "smri_vol_cdk_pcrh",
    "smri_vol_cdk_rracaterh", "smri_vol_cdk_rrmdfrrh", "smri_vol_cdk_sufrrh",
    "smri_vol_cdk_suplrh", "smri_vol_cdk_sutmrh", "smri_vol_cdk_smrh",
    "smri_vol_cdk_frpolerh", "smri_vol_cdk_tmpolerh", "smri_vol_cdk_trvtmrh",
    "smri_vol_cdk_insularh"
  ),
  region_name = c(
    "Left-Cerebral_Cortex", "Right-Cerebral-Cortex", "Total-Cerebral-Cortex",
    "Total_Cerebral-WM", "Total_Cerebellum-WM", "Total-Cerebellum-Cortex",
    
    "Left-Cerebral-White-Matter", "Left-Lateral-Ventricle", "Left-Inf-Lat-Vent",
    "Left-Cerebellum-White-Matter", "Left-Cerebellum-Cortex", "Left-Thalamus-Proper",
    "Left-Caudate", "Left-Putamen", "Left-Pallidum",
    "x3rd-ventricle", "x4th-ventricle", "Brain-Stem",
    "Left-Hippocampus", "Left-Amygdala", "Cerebrospinal-Fluid",
    "Left-Lesion", "Left-Accumbens-Area", "Left-VentralDC",
    
   "Right-Cerebral-White-Matter", "Right-Lateral-Ventricle", "Right-Inf-Lat-Vent",
    "Right-Cerebellum-White-Matter", "Right-Cerebellum-Cortex", "Right-Thalamus-Proper",
    "Right-Caudate", "Right-Putamen", "Right-Pallidum",
    "Right-Hippocampus", "Right-Amygdala",
    "Right-Lesion", "Right-Accumbens-Area", "Right-VentralDC",
    
    "WM-Hypointensities","Left-WM-Hypointensities", "Right-WM-Hypointensities",
   
   "CC_Posterior","CC_Mid_Posterior", "CC_Central", "CC_Mid_Anterior",
    "CC_Anterior", 
   
   "WholeBrain", "LatVentricles","AllVentricles", "IntracranialVolume", 
   "SupratentorialVolume", "SubcorticalGrayVolume",
   
     "lh_bankssts", "lh_caudalanteriorcingulate", "lh_caudalmiddlefrontal",
    "lh_cuneus", "lh_entorhinal", "lh_fusiform",
    "lh_inferiorparietal", "lh_inferiortemporal", "lh_isthmuscingulate",
    "lh_lateraloccipital", "lh_lateralorbitofrontal", "lh_lingual",
    "lh_medialorbitofrontal", "lh_middletemporal", "lh_parahippocampal",
    "lh_paracentral", "lh_parsopercularis", "lh_parsorbitalis",
    "lh_parstriangularis", "lh_pericalcarine", "lh_postcentral",
    "lh_posteriorcingulate", "lh_precentral", "lh_precuneus", "lh_rostralanteriorcingulate",
    "lh_rostralmiddlefrontal", "lh_superiorfrontal", "lh_superiorparietal",
    "lh_superiortemporal", "lh_supramarginal", "lh_frontalpole",
    "lh_temporalpole", "lh_transversetemporal", "lh_insula",
   
     "rh_bankssts", "rh_caudalanteriorcingulate", "rh_caudalmiddlefrontal",
    "rh_cuneus", "rh_entorhinal", "rh_fusiform",
    "rh_inferiorparietal", "rh_inferiortemporal", "rh_isthmuscingulate",
    "rh_lateraloccipital", "rh_lateralorbitofrontal", "rh_lingual",
    "rh_medialorbitofrontal", "rh_middletemporal", "rh_parahippocampal",
    "rh_paracentral", "rh_parsopercularis", "rh_parsorbitalis",
    "rh_parstriangularis", "rh_pericalcarine", "rh_postcentral",
    "rh_posteriorcingulate", "rh_precentral", "rh_precuneus", "rh_rostralanteriorcingulate",
    "rh_rostralmiddlefrontal", "rh_superiorfrontal", "rh_superiorparietal",
    "rh_superiortemporal", "rh_supramarginal", "rh_frontalpole",
    "rh_temporalpole", "rh_transversetemporal", "rh_insula"
    )
)

phenotypes_for_MC <- c(
    "smri_vol_scs_ltventriclelh", "smri_vol_scs_inflatventlh",
    "smri_vol_scs_crbwmatterlh", "smri_vol_scs_crbcortexlh", "smri_vol_scs_tplh",
    "smri_vol_scs_caudatelh", "smri_vol_scs_putamenlh", "smri_vol_scs_pallidumlh",
    "smri_vol_scs_3rdventricle", "smri_vol_scs_4thventricle", "smri_vol_scs_bstem",
    "smri_vol_scs_hpuslh", "smri_vol_scs_amygdalalh", "smri_vol_scs_aal", "smri_vol_scs_vedclh",
    
    "smri_vol_scs_ltventriclerh", "smri_vol_scs_inflatventrh",
    "smri_vol_scs_crbwmatterrh", "smri_vol_scs_crbcortexrh", "smri_vol_scs_tprh",
    "smri_vol_scs_caudaterh", "smri_vol_scs_putamenrh", "smri_vol_scs_pallidumrh",
    "smri_vol_scs_hpusrh", "smri_vol_scs_amygdalarh",
    "smri_vol_scs_aar", "smri_vol_scs_vedcrh", 
    
    "smri_vol_scs_ccps",
    "smri_vol_scs_ccmidps", "smri_vol_scs_ccct", "smri_vol_scs_ccmidat",
    "smri_vol_scs_ccat",
    
    "smri_vol_cdk_banksstslh", "smri_vol_cdk_cdacatelh", "smri_vol_cdk_cdmdfrlh",
    "smri_vol_cdk_cuneuslh", "smri_vol_cdk_ehinallh", "smri_vol_cdk_fusiformlh",
    "smri_vol_cdk_ifpllh", "smri_vol_cdk_iftmlh", "smri_vol_cdk_ihcatelh",
    "smri_vol_cdk_locclh", "smri_vol_cdk_lobfrlh", "smri_vol_cdk_linguallh",
    "smri_vol_cdk_mobfrlh", "smri_vol_cdk_mdtmlh", "smri_vol_cdk_parahpallh",
    "smri_vol_cdk_paracnlh", "smri_vol_cdk_parsopclh", "smri_vol_cdk_parsobislh",
    "smri_vol_cdk_parstgrislh", "smri_vol_cdk_pericclh", "smri_vol_cdk_postcnlh",
    "smri_vol_cdk_ptcatelh", "smri_vol_cdk_precnlh", "smri_vol_cdk_pclh",
    "smri_vol_cdk_rracatelh", "smri_vol_cdk_rrmdfrlh", "smri_vol_cdk_sufrlh",
    "smri_vol_cdk_supllh", "smri_vol_cdk_sutmlh", "smri_vol_cdk_smlh",
    "smri_vol_cdk_frpolelh", "smri_vol_cdk_tmpolelh", "smri_vol_cdk_trvtmlh",
    "smri_vol_cdk_insulalh",
    
     "smri_vol_cdk_banksstsrh", "smri_vol_cdk_cdacaterh", "smri_vol_cdk_cdmdfrrh",
    "smri_vol_cdk_cuneusrh", "smri_vol_cdk_ehinalrh", "smri_vol_cdk_fusiformrh",
    "smri_vol_cdk_ifplrh", "smri_vol_cdk_iftmrh", "smri_vol_cdk_ihcaterh",
    "smri_vol_cdk_loccrh", "smri_vol_cdk_lobfrrh", "smri_vol_cdk_lingualrh",
    "smri_vol_cdk_mobfrrh", "smri_vol_cdk_mdtmrh", "smri_vol_cdk_parahpalrh",
    "smri_vol_cdk_paracnrh", "smri_vol_cdk_parsopcrh", "smri_vol_cdk_parsobisrh",
    "smri_vol_cdk_parstgrisrh", "smri_vol_cdk_periccrh", "smri_vol_cdk_postcnrh",
    "smri_vol_cdk_ptcaterh", "smri_vol_cdk_precnrh", "smri_vol_cdk_pcrh",
    "smri_vol_cdk_rracaterh", "smri_vol_cdk_rrmdfrrh", "smri_vol_cdk_sufrrh",
    "smri_vol_cdk_suplrh", "smri_vol_cdk_sutmrh", "smri_vol_cdk_smrh",
    "smri_vol_cdk_frpolerh", "smri_vol_cdk_tmpolerh", "smri_vol_cdk_trvtmrh",
    "smri_vol_cdk_insularh"
  )

global_phenotypes <- c("smri_vol_cdk_total", "totalWM_cb", "smri_vol_scs_allventricles", "smri_vol_scs_subcorticalgv")

all_global_phenotypes <- c("smri_vol_cdk_total", "totalWM_cb", "smri_vol_scs_allventricles", "smri_vol_scs_subcorticalgv", "smri_vol_cdk_totallh", "smri_vol_cdk_totalrh", "totalWM_crb", "totalGM_crb", "smri_vol_scs_cbwmatterlh", "smri_vol_scs_crbwmatterlh", "smri_vol_scs_crbcortexlh","smri_vol_scs_cbwmatterrh", "smri_vol_scs_crbwmatterrh","smri_vol_scs_crbcortexrh", "smri_vol_scs_wholeb", "smri_vol_scs_latventricles", "smri_vol_scs_intracranialv", "smri_vol_scs_suprateialv")

phenotypes_to_fit <- c(phenotypes_for_MC, global_phenotypes, "smri_vol_scs_wholeb", "smri_vol_scs_intracranialv", "totalWM_crb", "totalGM_crb", "smri_vol_scs_csf", "smri_vol_cdk_totallh", "smri_vol_cdk_totalrh", "smri_vol_scs_cbwmatterlh", "smri_vol_scs_cbwmatterrh")

pull <- function(data, var) {
  if (var %in% names(data)) data[[var]] else NA
}
```


#Generate Models
```{r}
if (generate == TRUE){
family_set <- c("GG")
phenotype_set <- phenotypes_to_fit
out_folder <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/gamlss_models_long_vol_2.0/"

terms_mu = c("~1", "age", "sex", "random(site)", "interscan_months_t1t2", "age:sex")
terms_sigma = c("~1", "age", "sex")
terms_nu = c("~1", "age", "sex")
terms_tau = c("~1")

generate_gamlss_models(phenotype_set, 
                                   out_folder, 
                                   family_set = family_set,
                                   mu_terms = terms_mu,
                                   sigma_terms = terms_sigma,
                                   nu_terms = terms_nu,
                                   tau_terms = terms_tau,
                                   long = TRUE,
                                   n_crit = 200)
}
```
##Commands to fit the models
Run fits submitting an array job.
split 1
sbatch array_script_submit.sh submit_gamlss_fits_ABCD_ONE.sh /mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/gamlss_models_long_vol_2.0 "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/gamlss_models_long_vol_2.0/" "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/CSV/process_tables_5.1_2.0/abcd_long_Onetrain2025-06-16.csv" "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/gamlss_fits_long_vol_splitONE_2.0/" "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/gamlss_fits_long_vol_splitONE_2.0/gamlss_fits_stats" "long"

split 2
sbatch array_script_submit.sh submit_gamlss_fits_ABCD_TWO.sh /mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/gamlss_models_long_vol_2.0 "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/gamlss_models_long_vol_2.0/" "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/CSV/process_tables_5.1_2.0/abcd_long_Twotrain2025-06-16.csv" "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/gamlss_fits_long_vol_splitTWO_2.0/" "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/gamlss_fits_long_vol_splitTWO_2.0/gamlss_fits_stats" "long"


#Data Setup
##Define files
```{r}
#file paths
splitOne_folder <- paste0(fits_path, "gamlss_fits_long_vol_splitONE_2.0/")
splitTwo_folder <- paste0(fits_path, "gamlss_fits_long_vol_splitTWO_2.0/")

sOne_fits_stats_filename <- paste0(splitOne_folder, "gamlss_fits_stats_all.csv")
sTwo_fits_stats_filename <- paste0(splitTwo_folder, "gamlss_fits_stats_all.csv")

splitOne_train_filename <- paste0(data_path, "abcd_wide_Onetrain2025-06-17.csv")
splitTwo_train_filename <- paste0(data_path, "abcd_wide_Twotrain2025-06-17.csv")
splitOne_test_filename <- paste0(data_path, "abcd_wide_Onetest2025-06-17.csv")
splitTwo_test_filename <- paste0(data_path, "abcd_wide_Twotest2025-06-17.csv")
```
## Load data frames and clean them up
```{r}
#rename variables
rename_vec <- c(age = "interview_age.t2", sex = "sex_baseline", site = "site_id_l.t2")

sOne_train_df <- read.csv(splitOne_train_filename)  %>% select(-1) %>% rename(., all_of(rename_vec))
sOne_train_df$sex <- as.factor(sOne_train_df$sex)
sOne_train_df$site <- as.factor(sOne_train_df$site)
sOne_train_df$age <- as.numeric(sOne_train_df$age)

sTwo_train_df <- read.csv(splitTwo_train_filename)  %>% select(-1) %>% rename(., all_of(rename_vec))
sTwo_train_df$sex <- as.factor(sTwo_train_df$sex)
sTwo_train_df$site <- as.factor(sTwo_train_df$site)
sTwo_train_df$age <- as.numeric(sTwo_train_df$age)

sOne_test_df <- read.csv(splitOne_test_filename)  %>% select(-1) %>% rename(., all_of(rename_vec))
sOne_test_df$sex <- as.factor(sOne_test_df$sex)
sOne_test_df$site <- as.factor(sOne_test_df$site)
sOne_test_df$age <- as.numeric(sOne_test_df$age)

sTwo_test_df <- read.csv(splitTwo_test_filename)  %>% select(-1) %>% rename(., all_of(rename_vec))
sTwo_test_df$sex <- as.factor(sTwo_test_df$sex)
sTwo_test_df$site <- as.factor(sTwo_test_df$site)
sTwo_test_df$age <- as.numeric(sTwo_test_df$age)
```
#Fits
##load all fits, process, and save
Make a table of error messages, warnings etc
Phenotype set should not have included wmhint, those models gave errors: all wmhint models are deleted.
```{r}
all_fits_filename <- paste0(fits_path, "long_fits.csv")
if(save_fits == TRUE){
sOne_fits <- inspect_model_fits(splitOne_folder, sOne_fits_stats_filename)
sOne_fits$label <- name_mapping$region_name[match(sub(".t2", "", sOne_fits$phenotype), name_mapping$abbreviation)]
#remove models that have age:sex interaction without age and sex main effects
sOne_fits <- sOne_fits %>%
  filter(!(
    str_detect(text, "age:sex") & 
      !(str_detect(text, "\\+ age \\+") &
        str_detect(text, "\\+ sex \\+")))
  )

sTwo_fits <- inspect_model_fits(splitTwo_folder, sTwo_fits_stats_filename)
sTwo_fits$label <- name_mapping$region_name[match(sub(".t2", "", sTwo_fits$phenotype), name_mapping$abbreviation)]
#remove models that have age:sex interaction without age and sex main effects
sTwo_fits <- sTwo_fits %>%
  filter(!(
    str_detect(text, "age:sex") & 
      !(str_detect(text, "\\+ age \\+") &
        str_detect(text, "\\+ sex \\+")))
  )


sOne_fits$split_no <- rep("1", dim(sOne_fits)[1])
sTwo_fits$split_no <- rep("2", dim(sTwo_fits)[1])

all_fits <- rbind(sOne_fits, sTwo_fits)

write.csv(all_fits, all_fits_filename)
}
```
##inspect errors
```{r}
#load saved fits if saved.
if(save_fits == FALSE) {
  all_fits <- read.csv(all_fits_filename) %>% select(-1)
  sOne_fits <- all_fits %>% filter(split_no == 1)
  sTwo_fits <- all_fits %>% filter(split_no == 2)
}
#split One fits
 sOne_fits %>% filter(error == 1) %>% select(error_text)
#distribution of errors among phenotypes
table(sOne_fits %>% filter(error == 1) %>% select(phenotype))

#distribution of errors among model terms
table(sOne_fits %>% filter(error == 1) %>% select(mu_formula))
table(sOne_fits %>% filter(error == 1) %>% select(sigma_formula))
table(sOne_fits %>% filter(error == 1) %>% select(nu_formula))

#split Two fits
 sTwo_fits %>% filter(error == 1) %>% select(error_text)
 
#distribution of errors among phenotypes
table(sTwo_fits %>% filter(error == 1) %>% select(phenotype))

#distribution of errors among model terms
table(sTwo_fits %>% filter(error == 1) %>% select(mu_formula))
table(sTwo_fits %>% filter(error == 1) %>% select(sigma_formula))
table(sTwo_fits %>% filter(error == 1) %>% select(nu_formula))
 
```
##inspect convergence
```{r}
#split One fits
sum(sOne_fits$convergence_warn_end == 1)
table(sOne_fits %>% filter(convergence_warn_end == 1) %>% select(phenotype))
#split Two fits
sum(sTwo_fits$convergence_warn_end == 1)
table(sTwo_fits %>% filter(convergence_warn_end == 1) %>% select(phenotype))
```
##load saved fits and get the models with the lowest BIC within each phenotype
```{r}
#load saved fits if saved previously
if(save_fits == FALSE) {
  all_fits <- read.csv(all_fits_filename) %>% select(-1)
  sOne_fits <- all_fits %>% filter(split_no == 1)
  sTwo_fits <- all_fits %>% filter(split_no == 2)
}

sOne_best <- sOne_fits %>%
  filter(convergence_warn_end == 0) %>%
  group_by(phenotype) %>%
 slice_min(BIC, n = 1)

sTwo_best <- sTwo_fits %>%
  filter(convergence_warn_end == 0) %>%
  group_by(phenotype) %>%
  slice_min(BIC, n = 1)
```

#Inspect Best Models - by BIC
##get table of which vars are chosen for which phenotype & which moment
```{r}
#terms
terms_mu = c("age", "sex", "age:sex", "interscan_months_t1t2")
terms_sigmanu = c("age", "sex")
moments = c("mu", "sigma", "nu")
mu_colnames = paste(terms_mu, "mu", sep = ".")
sigmanu_colnames = apply(expand.grid(terms_sigmanu, c("sigma", "nu")), 1, paste, collapse=".")
colnames = c(mu_colnames, sigmanu_colnames)

#sOne
sOne_best_terms <- as.data.frame(matrix(nrow = dim(sOne_best)[1], ncol=length(colnames)))
names(sOne_best_terms) <- colnames
rownames(sOne_best_terms) <- sOne_best$phenotype
  for(i in 1:length(terms_mu)) {
      sOne_best_terms[,(i)] <- grepl(terms_mu[i], sOne_best$mu_formula)
  }
  for(i in 1:length(terms_sigmanu)) {
      sOne_best_terms[,(i+length(terms_mu))] <- grepl(terms_sigmanu[i], sOne_best$sigma_formula)
      sOne_best_terms[,(i+length(terms_mu)+length(terms_sigmanu))] <- grepl(terms_sigmanu[i], sOne_best$nu_formula)
  }

sOne_best_terms$label <- name_mapping$region_name[match(sub(".t2", "", rownames(sOne_best_terms)), name_mapping$abbreviation)]

#sTwo
sTwo_best_terms <- as.data.frame(matrix(nrow = dim(sTwo_best)[1], ncol=length(colnames)))
names(sTwo_best_terms) <- colnames
rownames(sTwo_best_terms) <- sTwo_best$phenotype
  for(i in 1:length(terms_mu)) {
      sTwo_best_terms[,(i)] <- grepl(terms_mu[i], sTwo_best$mu_formula)
  }
  for(i in 1:length(terms_sigmanu)) {
      sTwo_best_terms[,(i+length(terms_mu))] <- grepl(terms_sigmanu[i], sTwo_best$sigma_formula)
      sTwo_best_terms[,(i+length(terms_mu)+length(terms_sigmanu))] <- grepl(terms_sigmanu[i], sTwo_best$nu_formula)
  }

sTwo_best_terms$label <- name_mapping$region_name[match(sub(".t2", "", rownames(sTwo_best_terms)), name_mapping$abbreviation)]
```
##inspect distribution of terms in "BEST" models
###sOne
```{r}
#sOne
sum(sOne_best_terms$age.mu)
sum(sOne_best_terms$sex.mu)
sum(sOne_best_terms$`age:sex.mu`)
sum(sOne_best_terms$interscan_months_t1t2.mu)

sum(sOne_best_terms$age.sigma)
sum(sOne_best_terms$sex.sigma)

sum(sOne_best_terms$age.nu)
sum(sOne_best_terms$sex.nu)

#visualize regions based on terms in the model
##merge with atlas data
plot_data_dkt <- merge(sOne_best_terms, ggseg::dk, by = "label")
plot_data_aseg <- merge(sOne_best_terms, ggseg::aseg, by = "label")

#Regions not matched in atlases
sOne_best_terms[!(sOne_best_terms$label %in% plot_data_dkt$label) & !(sOne_best_terms$label %in% plot_data_aseg$label),] %>% print(., row.names = FALSE)

(long.sOne_agemu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age.mu)) +
  theme_void()+
  labs(title = "Age in Mu"))

(long.sOne_agemu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = age.mu)) +
  theme_void() +
  labs(title = "Age in Mu"))

(long.sOne_sexmu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex.mu)) +
  theme_void()+
  labs(title = "Sex in Mu"))

(long.sOne_sexmu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = sex.mu)) +
  theme_void() +
  labs(title = "Sex in Mu"))

(long.sOne_agesexmu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = `age:sex.mu`)) +
  theme_void()+
  labs(title = "AgeSex in Mu"))

(long.sOne_agesexmu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = `age:sex.mu`)) +
  theme_void() +
  labs(title = "AgeSex in Mu"))

(long.sOne_interscanmu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = interscan_months_t1t2.mu)) +
  theme_void()+
  labs(title = "Interscan in Mu"))

(long.sOne_interscanmu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = interscan_months_t1t2.mu)) +
  theme_void() +
  labs(title = "Interscan in Mu"))


(long.sOne_agesigma_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age.sigma)) +
  theme_void()+
  labs(title = "Age in Sigma"))

(long.sOne_agesigma_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = age.sigma)) +
  theme_void() +
  labs(title = "Age in Sigma"))

(long.sOne_sexsigma_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex.sigma)) +
  theme_void()+
  labs(title = "Sex in Sigma"))

(long.sOne_sexsigma_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = sex.sigma)) +
  theme_void() +
  labs(title = "Sex in Sigma"))

(long.sOne_agenu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age.nu)) +
  theme_void()+
  labs(title = "Age in Nu"))

(long.sOne_agenu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = age.nu)) +
  theme_void() +
  labs(title = "Age in Nu"))

(long.sOne_sexnu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex.nu)) +
  theme_void()+
  labs(title = "Sex in Nu"))

(long.sOne_sexnu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = sex.nu)) +
  theme_void() +
  labs(title = "Sex in Nu"))

```
###sTwo
```{r}
#sTwo
sum(sTwo_best_terms$age.mu)
sum(sTwo_best_terms$sex.mu)
sum(sTwo_best_terms$`age:sex.mu`)
sum(sTwo_best_terms$interscan_months_t1t2.mu)

sum(sTwo_best_terms$age.sigma)
sum(sTwo_best_terms$sex.sigma)

sum(sTwo_best_terms$age.nu)
sum(sTwo_best_terms$sex.nu)

#visualize regions based on terms in the model
##merge with atlas data
plot_data_dkt <- merge(sTwo_best_terms, ggseg::dk, by = "label")
plot_data_aseg <- merge(sTwo_best_terms, ggseg::aseg, by = "label")

#Regions not matched in atlases
sTwo_best_terms[!(sTwo_best_terms$label %in% plot_data_dkt$label) & !(sTwo_best_terms$label %in% plot_data_aseg$label),] %>% print(., row.names = FALSE)

(long.sTwo_agemu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age.mu)) +
  theme_void()+
  labs(title = "Age in Mu"))

(long.sTwo_agemu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = age.mu)) +
  theme_void() +
  labs(title = "Age in Mu"))

(long.sTwo_sexmu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex.mu)) +
  theme_void()+
  labs(title = "Sex in Mu"))

(long.sTwo_sexmu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = sex.mu)) +
  theme_void() +
  labs(title = "Sex in Mu"))

(long.sTwo_agesexmu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = `age:sex.mu`)) +
  theme_void()+
  labs(title = "AgeSex in Mu"))

(long.sTwo_agesexmu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = `age:sex.mu`)) +
  theme_void() +
  labs(title = "AgeSex in Mu"))

(long.sTwo_interscanmu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = interscan_months_t1t2.mu)) +
  theme_void()+
  labs(title = "Interscan in Mu"))

(long.sTwo_interscanmu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = interscan_months_t1t2.mu)) +
  theme_void() +
  labs(title = "Interscan in Mu"))


(long.sTwo_agesigma_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age.sigma)) +
  theme_void()+
  labs(title = "Age in Sigma"))

(long.sTwo_agesigma_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = age.sigma)) +
  theme_void() +
  labs(title = "Age in Sigma"))

(long.sTwo_sexsigma_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex.sigma)) +
  theme_void()+
  labs(title = "Sex in Sigma"))

(long.sTwo_sexsigma_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = sex.sigma)) +
  theme_void() +
  labs(title = "Sex in Sigma"))

(long.sTwo_agenu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age.nu)) +
  theme_void()+
  labs(title = "Age in Nu"))

(long.sTwo_agenu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = age.nu)) +
  theme_void() +
  labs(title = "Age in Nu"))

(long.sTwo_sexnu_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex.nu)) +
  theme_void()+
  labs(title = "Sex in Nu"))

(long.sTwo_sexnu_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = sex.nu)) +
  theme_void() +
  labs(title = "Sex in Nu"))
```
#Load best model
```{r}
sOne_best_list <- paste0(splitOne_folder,"gamlssFIT_",sOne_best$model_index,".rds")
best_models_One <- lapply(sOne_best_list, readRDS)
sTwo_best_list <- paste0(splitTwo_folder,"gamlssFIT_",sTwo_best$model_index,".rds")
best_models_Two <- lapply(sTwo_best_list, readRDS)
```
#Diagnostics
##Worm plots
get worm plots for best BIC within each phenotype, for each split
get worm plots for best BIC within each phenotype, for each split
###sOne
####worm plots
```{r}
if(diagnostics == TRUE){
plot <- FALSE
sOne_plots_wp <- vector(mode = "list", length = dim(sOne_best)[1])
sOne_wp_table <- data.frame(matrix(NA, nrow = dim(sOne_best)[1], ncol = 5))
names(sOne_wp_table) <- c("n", "n_outer", "pcnt", "phenotype", "label")
for (i  in 1:length(best_models_One)) {
  #print(deparse(model$mu.formula))
  model <- best_models_One[[i]]
  phenotype <- as.character(model$mu.terms[[2]])
  title <- name_mapping$region_name[match(sub(".t2", "", phenotype), name_mapping$abbreviation)]
      sOne_plots_wp[[i]] <- wp.taki(model, ylim.worm = 0.5, xlim.worm = 4) + labs(title = title)
      sOne_wp_table[i,1:3] <- wp.taki_EK(model, ylim.worm = 0.5, xlim.worm = 4, plot = FALSE) %>% select(all_of(c("n", "n_outer", "pcnt")))
      sOne_wp_table[i,4] <- phenotype
      sOne_wp_table[i,5] <- title
    }

if (plot == TRUE){
n <- length(sOne_plots_wp)
nCol <- min(c(floor(sqrt(n)), 4))
nRow <- min(ceiling(n/nCol), 4)
breaks <- c(seq(from = 0, to = n , by = (nCol*nRow)), n)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(sOne_plots_wp[start:end], ncol=nCol))
}
sum(sOne_wp_table$pcnt > 0.05)
sOne_wp_toplot <- sOne_plots_wp[which(sOne_wp_table$pcnt > 0.05)]

n <- length(sOne_wp_toplot)
nCol <- min(c(floor(sqrt(n)), 4))
nRow <- min(ceiling(n/nCol), 4)
breaks <- c(seq(from = 0, to = n , by = (nCol*nRow)), n)
if(breaks[length(breaks)] == breaks[length(breaks)-1]){
  breaks <- breaks[-length(breaks)]
}
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(sOne_wp_toplot[start:end], ncol=nCol))
}
}
}
```
####re-select best model considering wormplot fit for each phenotype that has >5% points outside for the best BIC
```{r}
sOne_refit_phenotypes <- sOne_wp_table %>% filter(pcnt > 0.05) %>% select(label)

sOne_best %>% filter(label %in% sOne_refit_phenotypes$label) %>% select(c(mu_formula))
sOne_best %>% filter(label %in% sOne_refit_phenotypes$label) %>% select(c(sigma_formula))
sOne_best %>% filter(label %in% sOne_refit_phenotypes$label) %>% select(c(nu_formula))

sOne_refit_indexes <- sOne_fits %>% filter(label %in% sOne_refit_phenotypes$label) %>% select(model_index)

sOne_refit_list <- paste0(splitOne_folder,"gamlssFIT_",sOne_refit_indexes$model_index,".rds")
refit_models_One <- lapply(sOne_refit_list, readRDS)

#inspect worm plots of the other fitted models for the above phenotypes
plot <- FALSE
sOne_plots_refit_wp <- vector(mode = "list", length = length(sOne_refit_list))
sOne_wp_refit_table <- data.frame(matrix(NA, nrow = dim(sOne_refit_indexes)[1], ncol = 6))
names(sOne_wp_refit_table) <- c("n", "n_outer", "pcnt", "phenotype", "label", "model_index")
for (i  in 1:length(refit_models_One)) {
  #print(deparse(model$mu.formula))
  model <- refit_models_One[[i]]
  phenotype <- as.character(model$mu.terms[[2]])
  title <- name_mapping$region_name[match(sub(".t2", "", phenotype), name_mapping$abbreviation)]
      sOne_plots_refit_wp[[i]] <- wp.taki(model, ylim.worm = 0.5, xlim.worm = 4) + labs(title = title)
      sOne_wp_refit_table[i,1:3] <- wp.taki_EK(model, ylim.worm = 0.5, xlim.worm = 4, plot = FALSE) %>% select(all_of(c("n", "n_outer", "pcnt")))
      sOne_wp_refit_table[i,4] <- phenotype
      sOne_wp_refit_table[i,5] <- title
      sOne_wp_refit_table[i,6] <- sOne_refit_indexes$model_index[i]
    }

if (plot == TRUE){
n <- length(sOne_plots_refit_wp)
nCol <- min(c(floor(sqrt(n)), 4))
nRow <- min(ceiling(n/nCol), 4)
breaks <- c(seq(from = 0, to = n , by = (nCol*nRow)), n)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(sOne_plots_refit_wp[start:end], ncol=nCol))
}
}


sOne_wp_refit_table <- merge(sOne_wp_refit_table, sOne_fits %>% select(c("model_index", "BIC")), all.x = TRUE, all.y = FALSE, by = "model_index")

#inspect distribution of wormplots "out points"
aggregate(pcnt ~ phenotype, data = sOne_wp_refit_table, summary)

sOne_fits <- sOne_fits %>% group_by(phenotype) %>%
  mutate(logminBICdiff = log(BIC-min(BIC))) %>%
  mutate( minBICdiff = BIC-min(BIC)) %>%
  mutate( BICzscore = scale(BIC)) %>%
  mutate(BICzscorediff = BICzscore-min(BICzscore)) %>%
  ungroup()

sOne_wp_refit_table <- merge(sOne_wp_refit_table, sOne_fits %>% select(c("model_index", "logminBICdiff", "BIC", "minBICdiff", "BICzscorediff")), all.x = TRUE, all.y = FALSE, by = "model_index")

sOne_wp_refit_table <- sOne_wp_refit_table %>% group_by(phenotype) %>%
  mutate(logminNoutdiff = log(pcnt-min(pcnt))) %>%
  mutate(minNoutdiff = pcnt - min(pcnt)) %>%
  mutate( Noutzscore = scale(pcnt)) %>%
  mutate(Noutzscorediff = Noutzscore-min(Noutzscore)) %>%
  ungroup()

sOne_wp_refit_table$wpBIC_metric <- 0.45*sOne_wp_refit_table$BICzscorediff + 0.55*sOne_wp_refit_table$Noutzscorediff

sOne_refit_best <- sOne_wp_refit_table %>%
  group_by(phenotype) %>%
  slice_min(wpBIC_metric, n = 1)

sOne_fits$best <- ifelse(sOne_fits$model_index %in% sOne_best$model_index | sOne_fits$model_index %in% sOne_refit_best$model_index, TRUE, FALSE)

sOne_fits %>% filter(label %in% sOne_refit_phenotypes$label) %>%
  ggplot(., aes(x = logminBICdiff, y = as.character(model_index), fill = best)) + 
  geom_histogram(stat = "identity") + 
  facet_wrap(~label, scale = "free") +
  ggtitle("BIC for refit") + labs(x = "scaled log BIC", y = "model index")

sOne_wp_refit_table$best <- sOne_wp_refit_table$model_index %in% sOne_refit_best$model_index

sOne_wp_refit_table %>% 
  ggplot(., aes(x = pcnt, y = as.character(model_index), fill = best)) + 
  geom_histogram(stat = "identity") + 
  facet_wrap(~label, scale = "free") +
  ggtitle("Nout for refit") + labs(x = "% out of bounds", y = "model index")


```
####adjust best Models based on above
```{r}
#add best model indices from the lowest BIC best models
sOne_final_best_model_index <- sOne_best %>% filter(!(label %in% sOne_refit_phenotypes$label)) %>% select(model_index)
#add best model indices from the BIC+wp combined metric
sOne_final_best_model_index <- c(sOne_final_best_model_index$model_index, sOne_refit_best$model_index)
#load all new best models
sOne_best_list <- paste0(splitOne_folder,"gamlssFIT_",sOne_final_best_model_index,".rds")
best_models_One <- lapply(sOne_best_list, readRDS)
```
###sTwo
####worm plots
```{r}
#diagnostics = TRUE
if(diagnostics == TRUE){
plot <- FALSE
sTwo_plots_wp <- vector(mode = "list", length = dim(sTwo_best)[1])
sTwo_wp_table <- data.frame(matrix(NA, nrow = dim(sTwo_best)[1], ncol = 5))
names(sTwo_wp_table) <- c("n", "n_outer", "pcnt", "phenotype", "label")
for (i  in 1:length(best_models_Two)) {
  #print(deparse(model$mu.formula))
  model <- best_models_Two[[i]]
  phenotype <- as.character(model$mu.terms[[2]])
  title <- name_mapping$region_name[match(sub(".t2", "", phenotype), name_mapping$abbreviation)]
      sTwo_plots_wp[[i]] <- wp.taki(model, ylim.worm = 0.5, xlim.worm = 4) + labs(title = title)
      sTwo_wp_table[i,1:3] <- wp.taki_EK(model, ylim.worm = 0.5, xlim.worm = 4, plot = FALSE) %>% select(all_of(c("n", "n_outer", "pcnt")))
      sTwo_wp_table[i,4] <- phenotype
      sTwo_wp_table[i,5] <- title
    }

if (plot == TRUE){
n <- length(sTwo_plots_wp)
nCol <- min(c(floor(sqrt(n)), 4))
nRow <- min(ceiling(n/nCol), 4)
breaks <- c(seq(from = 0, to = n , by = (nCol*nRow)), n)
for(k in 1:length(breaks)){
  start <- breaks[k] + 1
  end <- breaks[k+1]
  do.call("grid.arrange", c(sTwo_plots_wp[start:end], ncol=nCol))
}
}
sum(sTwo_wp_table$pcnt > 0.05)
sTwo_wp_toplot <- sTwo_plots_wp[which(sTwo_wp_table$pcnt > 0.05)]

n <- length(sTwo_wp_toplot)
nCol <- min(c(floor(sqrt(n)), 4))
nRow <- min(ceiling(n/nCol), 4)
breaks <- c(seq(from = 0, to = n , by = (nCol*nRow)), n)
if(breaks[length(breaks)] == breaks[length(breaks)-1]){
  breaks <- breaks[-length(breaks)]
}
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(sTwo_wp_toplot[start:end], ncol=nCol))
}
}
```
####re-select best model considering wormplot fit for each phenotype that has >5% points outside for the best BIC
```{r}
sTwo_refit_phenotypes <- sTwo_wp_table %>% filter(pcnt > 0.05) %>% select(label)

sTwo_best %>% filter(label %in% sTwo_refit_phenotypes$label) %>% select(c(mu_formula))
sTwo_best %>% filter(label %in% sTwo_refit_phenotypes$label) %>% select(c(sigma_formula))
sTwo_best %>% filter(label %in% sTwo_refit_phenotypes$label) %>% select(c(nu_formula))

sTwo_refit_indexes <- sTwo_fits %>% filter(label %in% sTwo_refit_phenotypes$label) %>% select(model_index)

sTwo_refit_list <- paste0(splitOne_folder,"gamlssFIT_",sTwo_refit_indexes$model_index,".rds")
refit_models_One <- lapply(sTwo_refit_list, readRDS)

#inspect worm plots of the other fitted models for the above phenotypes
plot <- FALSE
sTwo_plots_refit_wp <- vector(mode = "list", length = length(sTwo_refit_list))
sTwo_wp_refit_table <- data.frame(matrix(NA, nrow = dim(sTwo_refit_indexes)[1], ncol = 6))
names(sTwo_wp_refit_table) <- c("n", "n_outer", "pcnt", "phenotype", "label", "model_index")
for (i  in 1:length(refit_models_One)) {
  #print(deparse(model$mu.formula))
  model <- refit_models_One[[i]]
  phenotype <- as.character(model$mu.terms[[2]])
  title <- name_mapping$region_name[match(sub(".t2", "", phenotype), name_mapping$abbreviation)]
      sTwo_plots_refit_wp[[i]] <- wp.taki(model, ylim.worm = 0.5, xlim.worm = 4) + labs(title = title)
      sTwo_wp_refit_table[i,1:3] <- wp.taki_EK(model, ylim.worm = 0.5, xlim.worm = 4, plot = FALSE) %>% select(all_of(c("n", "n_outer", "pcnt")))
      sTwo_wp_refit_table[i,4] <- phenotype
      sTwo_wp_refit_table[i,5] <- title
      sTwo_wp_refit_table[i,6] <- sTwo_refit_indexes$model_index[i]
    }

if (plot == TRUE){
n <- length(sTwo_plots_refit_wp)
nCol <- min(c(floor(sqrt(n)), 4))
nRow <- min(ceiling(n/nCol), 4)
breaks <- c(seq(from = 0, to = n , by = (nCol*nRow)), n)
for(k in 1:length(breaks)){
  start <- breaks[k] + 1
  end <- breaks[k+1]
  do.call("grid.arrange", c(sTwo_plots_refit_wp[start:end], ncol=nCol))
}
}


sTwo_wp_refit_table <- merge(sTwo_wp_refit_table, sTwo_fits %>% select(c("model_index", "BIC")), all.x = TRUE, all.y = FALSE, by = "model_index")

#inspect distribution of wormplots "out points"
aggregate(pcnt ~ phenotype, data = sTwo_wp_refit_table, summary)

sTwo_fits <- sTwo_fits %>% group_by(phenotype) %>%
  mutate(logminBICdiff = log(BIC-min(BIC))) %>%
  mutate( minBICdiff = BIC-min(BIC)) %>%
  mutate( BICzscore = scale(BIC)) %>%
  mutate(BICzscorediff = BICzscore-min(BICzscore)) %>%
  ungroup()

sTwo_wp_refit_table <- merge(sTwo_wp_refit_table, sTwo_fits %>% select(c("model_index", "logminBICdiff", "BIC", "minBICdiff", "BICzscorediff")), all.x = TRUE, all.y = FALSE, by = "model_index")

sTwo_wp_refit_table <- sTwo_wp_refit_table %>% group_by(phenotype) %>%
  mutate(logminNoutdiff = log(pcnt-min(pcnt))) %>%
  mutate(minNoutdiff = pcnt - min(pcnt)) %>%
  mutate( Noutzscore = scale(pcnt)) %>%
  mutate(Noutzscorediff = Noutzscore-min(Noutzscore)) %>%
  ungroup()

sTwo_wp_refit_table$wpBIC_metric <- 0.45*sTwo_wp_refit_table$BICzscorediff + 0.55*sTwo_wp_refit_table$Noutzscorediff

sTwo_refit_best <- sTwo_wp_refit_table %>%
  group_by(phenotype) %>%
  slice_min(wpBIC_metric, n = 1)

sTwo_fits$best <- ifelse(sTwo_fits$model_index %in% sTwo_best$model_index | sTwo_fits$model_index %in% sTwo_refit_best$model_index, TRUE, FALSE)

sTwo_fits %>% filter(label %in% sTwo_refit_phenotypes$label) %>%
  ggplot(., aes(x = logminBICdiff, y = as.character(model_index), fill = best)) + 
  geom_histogram(stat = "identity") + 
  facet_wrap(~label, scale = "free") +
  ggtitle("BIC for refit") + labs(x = "scaled log BIC", y = "model index")

sTwo_wp_refit_table$best <- sTwo_wp_refit_table$model_index %in% sTwo_refit_best$model_index

sTwo_wp_refit_table %>% 
  ggplot(., aes(x = pcnt, y = as.character(model_index), fill = best)) + 
  geom_histogram(stat = "identity") + 
  facet_wrap(~label, scale = "free") +
  ggtitle("Nout for refit") + labs(x = "% out of bounds", y = "model index")


```
####adjust best Models based on above
```{r}
#add best model indices from the lowest BIC best models
sTwo_final_best_model_index <- sTwo_best %>% filter(!(label %in% sTwo_refit_phenotypes$label)) %>% select(model_index)
#add best model indices from the BIC+wp combined metric
sTwo_final_best_model_index <- c(sTwo_final_best_model_index$model_index, sTwo_refit_best$model_index)
#load all new best models
sTwo_best_list <- paste0(splitTwo_folder,"gamlssFIT_",sTwo_final_best_model_index,".rds")
best_models_Two <- lapply(sTwo_best_list, readRDS)
```
##cent-cdf
```{r}
if (diagnostics == TRUE){
print <- TRUE

centcdf_one <- setNames(data.frame(matrix(ncol = 3, nrow = dim(sOne_best)[1])), c("ph", "cdfcor_train", "cdfcor_test"))
centcdf_two <- setNames(data.frame(matrix(ncol = 3, nrow = dim(sTwo_best)[1])), c("ph", "cdfcor_train", "cdfcor_test"))
cdf_expected <- c(0.01, 0.05, 0.10, 0.25, 0.50, 0.75, 0.90, 0.95, 0.99)

for(i in 1:length(best_models_One)){
#define models from split one and split two
model_one <- best_models_One[[i]]
model_two <- best_models_Two[sapply(best_models_Two, function(x){any(grepl(as.character(model_one$mu.formula[[2]]),x$mu.formula))})][[1]]

title_one <- as.character(model_one$mu.formula[[2]])
title_two <- as.character(model_two$mu.formula[[2]])
stopifnot(title_one == title_two)
ph.t1 <- sub(".t2", ".t1", title_one)

#define original df from split one and two, with phenotypes from the current model
og_df_one <- sOne_train_df %>% select(c("age", "sex", "site", "interscan_months_t1t2", model_one$mu.formula[[2]], ph.t1))
og_df_two <- sTwo_train_df %>% select(c("age", "sex", "site", "interscan_months_t1t2", model_two$mu.formula[[2]], ph.t1))

#calculate predicted centiles distribution for sOne models, using sOne train dataset and sTwo test dataset
#save correlation with expected cdf values
model <- model_one 
out_train_one <- suppressMessages(cent_cdf(model_one, og_df_one))$data
#out_test_one <- suppressMessages(cent_cdf(model_one, sTwo_test_df))
centcdf_one$ph[i] <- title_one
centcdf_one$cdfcor_train[i] <- (cor.test(out_train_one$empirical, out_train_one$fitted)$estimate)
#centcdf_one$cdfcor_test[i] <- (cor.test(as.numeric(out_test_one[1,]), cdf_expected)$estimate)

#calculate predicted centiles distribution for sTwo models, using sTwo train dataset, and sOne test dataset
#save correlation with expected cdf values
model <- model_two
out_train_two <- suppressMessages(cent_cdf(model_two, og_df_two))$data
#out_test_two <- suppressMessages(cent_cdf(model_two, sOne_test_df))
centcdf_two$ph[i] <- title_two
centcdf_two$cdfcor_train[i] <-  (cor.test(out_train_two$empirical, out_train_two$fitted)$estimate)
#centcdf_two$cdfcor_test[i] <- (cor.test(as.numeric(out_test_two[1,]), cdf_expected)$estimate)


if (print == TRUE){
print(paste0("Correlation Coefficient Split One, Train Data ", title_one, centcdf_one$cdfcor_train[i]))
print(paste0("Correlation Coefficient Split Two, Train Data ", title_two, centcdf_two$cdfcor_train[i]))
}
}
}
```
#Predict Centile
predict centile scores for unseen split-half
```{r}
plot <- FALSE
print <- FALSE 

centiles_pred_one <- vector(mode = "list", length = dim(sOne_best)[1])
centiles_pred_two <- vector(mode = "list", length = dim(sTwo_best)[1])
splits_correlations <- setNames(data.frame(matrix(ncol = 4, nrow = dim(sOne_best)[1])), c("ph_one", "cor_est_one", "ph_two", "cor_est_two"))

for(i in 1:length(best_models_One)){
#define models from split one and split two
model_one <- best_models_One[[i]]
model_two <- best_models_Two[sapply(best_models_Two, function(x){any(grepl(as.character(model_one$mu.formula[[2]]),x$mu.formula))})][[1]]

title_one <- as.character(model_one$mu.formula[[2]])
title_two <- as.character(model_two$mu.formula[[2]])
stopifnot(title_one == title_two)
ph.t1 <- sub(".t2", ".t1", title_one)

#define original df from split one and two, with phenotypes from the current model
og_df_one <- sOne_train_df %>% select(c("age", "sex", "site", "interscan_months_t1t2", model_one$mu.formula[[2]], ph.t1))
og_df_two <- sTwo_train_df %>% select(c("age", "sex", "site", "interscan_months_t1t2", model_two$mu.formula[[2]], ph.t1))

#calculate predicted centiles on sOne and sTwo data using sOne model

model <- model_one
sOne_centiles_1 <- suppressMessages(pred_og_centile_EK(model_one, sOne_train_df))
sTwo_centiles_1 <- suppressMessages(pred_og_centile_EK(model_one, og_df_one, new.data = sTwo_train_df))
centiles_pred_two[[i]] <- suppressMessages(pred_og_centile_EK(model_one, og_df_one, new.data = sTwo_test_df))
names(centiles_pred_two)[i] <- paste(title_two, "centiles", sep = "_")
splits_correlations$ph_one[i] <- title_one

#calculate predicted centiles on sOne and sTwo data using sTwo model
model <- model_two
sOne_centiles_2 <- suppressMessages(pred_og_centile_EK(model_two, og_df_two, new.data = sOne_train_df))
sTwo_centiles_2 <- suppressMessages(pred_og_centile_EK(model_two, sTwo_train_df))
centiles_pred_one[[i]] <- suppressMessages(pred_og_centile_EK(model_two, og_df_two, new.data = sOne_test_df))
names(centiles_pred_one)[i] <- paste(title_one, "centiles", sep = "_")
splits_correlations$ph_two[i] <- title_two

if(plot == TRUE) {
  #Plotting sOne and sTwo from two predictions
  plot(sOne_centiles_1, sOne_centiles_2, main = paste0("split One: ", title_one) , ylab = "predictions from Model 1", xlab = "predictions from Model 2")
  abline(a = 0, b = 1, col = "red", lwd = 2, lty = 2)
  plot(sTwo_centiles_1, sTwo_centiles_2, main = paste0("split Two: ", title_two) , ylab = "predictions from Model 1", xlab = "predictions from Model 2")
  abline(a = 0, b = 1, col = "red", lwd = 2, lty = 2)
  }

#Correlation coefficients
cor_one <- cor.test(sOne_centiles_1, sOne_centiles_2)
cor_two <- cor.test(sTwo_centiles_1, sTwo_centiles_2)

#Save correlations in dataframe
splits_correlations$cor_est_one[i] <- cor_one$estimate
splits_correlations$cor_est_two[i] <- cor_two$estimate

if (print == TRUE){
print(paste0("Correlation Coefficient Split One, ", title_one, cor_one$estimate, " , ", cor_one$p.value))
print(paste0("Correlation Coefficient Split Two, ", title_two, cor_two$estimate, " , ", cor_two$p.value))
}
}
```
##Save predicted centiles
Save in single dataframe with columns for participant ID and split ID
```{r}
if(save_fits == TRUE){
output_folder <- paste0(fits_path, "centiles_calculated/")
if (!dir.exists(output_folder)) dir.create(output_folder)  
filename <- paste0(output_folder, "long.volume_centiles_", Sys.Date(), ".csv")

centiles_One <- cbind(sOne_test_df %>% select(src_subject_id), as.data.frame(centiles_pred_one))
centiles_One$split_ID <- rep("1", dim(centiles_One)[1])
                             
centiles_Two <- cbind(sTwo_test_df %>% select(src_subject_id), as.data.frame(centiles_pred_two))
centiles_Two$split_ID <- rep("2", dim(centiles_Two)[1])

centiles_all <- rbind(centiles_One, centiles_Two)

write.csv(centiles_all, filename)
}
```
#Centiles fit metrics
##test uniformity
```{r}
if(sum(grepl("centiles",names(sOne_test_df)))== 0) {
sOne_test_df <- cbind(sOne_test_df, as.data.frame(centiles_pred_one))
sTwo_test_df <- cbind(sTwo_test_df, as.data.frame(centiles_pred_two))
}
full_test_df <- rbind(sOne_test_df,sTwo_test_df)

#sOne
kstest_one <- as.data.frame(t(sapply(sOne_test_df[, grepl("centiles", names(sOne_test_df))], function(x) {
  #jitter data with small noise to avoid "ties" warning
  x <- x + runif(length(x), -1e-6, 1e-6)
  test <- ks.test(x, "punif", 0, 1)

  c(p_value = test$p.value[1], test = "uniformity")
})))

kstest_one$p_adj <- p.adjust(kstest_one$p_value, method = "BH")
kstest_one$label <- name_mapping$region_name[match(sub(".t2_centiles","",rownames(kstest_one)), name_mapping$abbreviation)]

kstest_one %>% filter(p_adj < 0.05) %>% select(label)

#sTwo
kstest_two <- as.data.frame(t(sapply(sTwo_test_df[, grepl("centiles", names(sTwo_test_df))], function(x) {
  #jitter data with small noise to avoid "ties" warning
  x <- x + runif(length(x), -1e-6, 1e-6)
  test <- ks.test(x, "punif", 0, 1)

  c(p_value = test$p.value[1], test = "uniformity")
})))

kstest_two$p_adj <- p.adjust(kstest_two$p_value, method = "BH")
m
```
#Save Parameter Table for Best Models
##sOne
```{r}
sOne_model_params_df <- data.frame(matrix(nrow = length(best_models_One), ncol = 20))
names(sOne_model_params_df) <- c("group","pheno", "family", "BIC", "mu_intercept","mu_pheno_t1", "mu_interscan", "mu_age", "mu_sexM", "mu_agesexM",
                                 "mu_randomsite_var", "mu_randomsite_std", "mu_randomsite_lambda",
                                 "sigma_intercept", "sigma_age", "sigma_sexM", 
                                 "nu_intercept", "nu_age", "nu_sexM", 
                                 "tau_intercept")

for (i in 1:length(best_models_One)){
  model = best_models_One[[i]]
  sOne_model_params_df$family[i] <- family(model)[[1]]
  sOne_model_params_df$BIC[i] <- BIC(model)
  phenotype <- sub(".t2","",as.character(formula(model)[[2]]))
  sOne_model_params_df$pheno[i] <- phenotype
  #Gather parameters
  ##mu
  sOne_model_params_df$mu_intercept[i] <- model$mu.coefficients[["(Intercept)"]]
  sOne_model_params_df$mu_pheno_t1[i] <- model$mu.coefficients[[2]]
  sOne_model_params_df$mu_interscan[i] <- pull(model$mu.coefficients, "interscan_months_t1t2")
  sOne_model_params_df$mu_age[i] <- pull(model$mu.coefficients, "age")
  sOne_model_params_df$mu_sexM[i] <- pull(model$mu.coefficients, "sexM")
  sOne_model_params_df$mu_agesexM[i] <- pull(model$mu.coefficients, "age:sexM")
  sOne_model_params_df$mu_randomsite_var[i] <- model$mu.coefSmo[[1]]$sigb2
  sOne_model_params_df$mu_randomsite_std[i] <- model$mu.coefSmo[[1]]$sigb
  sOne_model_params_df$mu_randomsite_lambda[i] <- model$mu.coefSmo[[1]]$lambda
  ##sigma
  sOne_model_params_df$sigma_intercept[i] <- model$sigma.coefficients[["(Intercept)"]]
  sOne_model_params_df$sigma_age[i] <- pull(model$sigma.coefficients, "age")
  sOne_model_params_df$sigma_sexM[i] <- pull(model$sigma.coefficients, "sexM")
  ##nu
  sOne_model_params_df$nu_intercept[i] <- model$nu.coefficients[["(Intercept)"]]
  sOne_model_params_df$nu_age[i] <- pull(model$nu.coefficients, "age")
  sOne_model_params_df$nu_sexM[i] <- pull(model$nu.coefficients, "sexM")
  ##tau
  sOne_model_params_df$tau_intercept[i] <- model$tau.coefficients[["(Intercept)"]]
}

sOne_model_params_df$group <- ifelse(sOne_model_params_df$pheno %in% all_global_phenotypes, "global", "region")

##save table as csv
filename <- "long.sOne.modelparams.csv"
if (!dir.exists(paste0(fits_path, "model_params_tables/"))) dir.create(paste0(fits_path, "model_params_tables/"))
write.csv(sOne_model_params_df, paste0(fits_path, "model_params_tables/", filename))
```
##sTwo
```{r}
sTwo_model_params_df <- data.frame(matrix(nrow = length(best_models_Two), ncol = 20))
names(sTwo_model_params_df) <- c("group","pheno", "family", "BIC", "mu_intercept","mu_pheno_t1", "mu_interscan", "mu_age", "mu_sexM", "mu_agesexM",
                                 "mu_randomsite_var", "mu_randomsite_std", "mu_randomsite_lambda",
                                 "sigma_intercept", "sigma_age", "sigma_sexM", 
                                 "nu_intercept", "nu_age", "nu_sexM", 
                                 "tau_intercept")

for (i in 1:length(best_models_Two)){
  model = best_models_Two[[i]]
  sTwo_model_params_df$family[i] <- family(model)[[1]]
  sTwo_model_params_df$BIC[i] <- BIC(model)
  phenotype <- sub(".t2","",as.character(formula(model)[[2]]))
  sTwo_model_params_df$pheno[i] <- phenotype
   #Gather parameters
  ##mu
  sTwo_model_params_df$mu_intercept[i] <- model$mu.coefficients[["(Intercept)"]]
  sTwo_model_params_df$mu_pheno_t1[i] <- model$mu.coefficients[[2]]
  sTwo_model_params_df$mu_interscan[i] <- pull(model$mu.coefficients, "interscan_months_t1t2")
  sTwo_model_params_df$mu_age[i] <- pull(model$mu.coefficients, "age")
  sTwo_model_params_df$mu_sexM[i] <- pull(model$mu.coefficients, "sexM")
  sTwo_model_params_df$mu_agesexM[i] <- pull(model$mu.coefficients, "age:sexM")
  sTwo_model_params_df$mu_randomsite_var[i] <- model$mu.coefSmo[[1]]$sigb2
  sTwo_model_params_df$mu_randomsite_std[i] <- model$mu.coefSmo[[1]]$sigb
  sTwo_model_params_df$mu_randomsite_lambda[i] <- model$mu.coefSmo[[1]]$lambda
  ##sigma
  sTwo_model_params_df$sigma_intercept[i] <- model$sigma.coefficients[["(Intercept)"]]
  sTwo_model_params_df$sigma_age[i] <- pull(model$sigma.coefficients, "age")
  sTwo_model_params_df$sigma_sexM[i] <- pull(model$sigma.coefficients, "sexM")
  ##nu
  sTwo_model_params_df$nu_intercept[i] <- model$nu.coefficients[["(Intercept)"]]
  sTwo_model_params_df$nu_age[i] <- pull(model$nu.coefficients, "age")
  sTwo_model_params_df$nu_sexM[i] <- pull(model$nu.coefficients, "sexM")
  ##tau
  sTwo_model_params_df$tau_intercept[i] <- model$tau.coefficients[["(Intercept)"]]
}

sTwo_model_params_df$group <- ifelse(sTwo_model_params_df$pheno %in% all_global_phenotypes, "global", "region")

##save table as csv
filename <- "long.sTwo.modelparams.csv"
if (!dir.exists(paste0(fits_path, "model_params_tables/"))) dir.create(paste0(fits_path, "model_params_tables/"))
write.csv(sTwo_model_params_df, paste0(fits_path, "model_params_tables/", filename))
```